Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-20 cav2017/cav2017.tex
revision intro synth section

%-------------------------------
\subsection{Digital implementation of feedback controller}
\label{ssec:quantizationerror}
%-------------------------------

\todo{HERE add a figure similar to Fig. \label{fig:observersystem} to explain digital features of controller.}

\todo{[The remainder of the section needs to be updated. ]}
The output and measurable of the plant is being sampled by an ADC, and since the full state of the
plant cannot be witnessed directly, the controller software includes an observer that 
estimates the state of the plant (see appendix for details on how such an observer
is constructed). In this work, we are assuming that all the states measurements are available for the state feedback purpose, and the state observer design may be disregarded.  
\todo{[Unclear what the remainder is for: ]}
We further presume that there are small tolerances in the system 
described in the following form
%
\begin{align*}
\mat{A} \in \mathcal{G}_A=\mat{S}\mat{J}\mat{S}^{-1} : \mat{J} \in \mathcal{G}_J
\end{align*}
\commentib{What is S?}

Where $\mathcal{G}_J$ is a collection of eigenvalue ranges that define the frequency response of the system.
We presume that the internal model of the plant is an ARMA model, modified by a
state transform that has no tolerances. Tolerances in the eigenvectors will manifest themselves
as tolerances in the canonical forms outside of the row/column of interests and complicate the 
evaluation. Whilst it is possible to address these with this methodology, we will focus on the case
above for simplicity.

@@

{\bf Assumption:} We consider the reference signal to be 0, whereas the initial states are non-deterministic.
\todo{Explain all the assumptions.}

%\todo{Integrate this paragraph better (Cristina):}
Our CEGIS technique with generalisation %abstraction refinement
relies on parametrising the target language, which
induces a lattice of progressively more expressive languages.  We
start by attempting to synthesise a program at the lowest point on
this lattice and increase the parameters of the language until we
reach a point at which the synthesis succeeds. As well as giving us an
automatic search procedure, this parametrisation greatly increases the
efficiency of our system since languages low down the lattice are very
easy to decide safety for (i.e. the validation oracle finds an answer
fast). If a program can be synthesised in a low-complexity language,
the whole procedure finishes much faster than if synthesis had been
attempted in a high-complexity language.

We investigate two different approaches of integrating solution generalisation 
with CEGIS:
\begin{itemize}
\item[1.] The first approach uses abstract acceleration to compute the set of reachable states, which enables it to perform abstraction refinement inside the CEGIS loop (see Sec~\ref{sec:CEGIS-abstraction-refinement}). 
\item[2.] The second approach relies on computing a 
completeness threshold $k$ for each loop such that it is sufficient to 
unroll the corresponding loop $k$ times. For this approach the 
generalisation step is given as an incremental
increase of precision and sampling rate (see Sec.~\ref{sec:CEGIS-precision-incrementation}).
\end{itemize}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{General architecture of the program synthesizer}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\label{synthesizer-general}
%
%% \red{[I would recommend to clarify the domain of definition of $\sigma$, particularly over the set where $x$ ranges over. 
%% (Be cognisant that control audience might be unfamiliar with the cegis architecture.) ]}
% 
The input specification provided to the program synthesizer is of the form
$\exists \vec{P} .  \forall \vec{x}.  \sigma(\vec{x}, \vec{P})$ where
$\vec{P}$ ranges over functions, $\vec{x}$ ranges over ground terms and
$\sigma$ is a quantifier-free formula.  We interpret the ground terms over
some finite domain $\mathcal{D}$.

The design of our synthesizer 
%is given in Figure~\ref{fig:CEGIS} and 
consists of two phases, an inductive synthesis phase and a validation phase,
%{\sc Synthesize} and {\sc Verify}, 
which interact via a finite set of test vectors {\sc inputs} that is updated incrementally. 
Given the aforementioned specification $\sigma$, the inductive synthesis
procedure
tries to find an existential witness $\vec{P}$ satisfying the specification
$\sigma(\vec{x}, \vec{P})$ for all $\vec{x}$ in {\sc inputs} (as opposed to
all $\vec{x} \in \mathcal{D}$).
%
If the synthesis phase succeeds in finding a witness~$\vec{P}$, this witness
is a candidate solution to the full synthesis formula.  We pass this
candidate solution to the validation phase, which checks whether it is a full
solution (i.e., $\vec{P}$ satisfies the specification $\sigma(\vec{x},
\vec{P})$ for all $\vec{x}\in\mathcal{D}$).
%
%determines whether
%it does satisfy the specification on all inputs by checking
%satisfiability of the following verification formula:
%
%$\exists \vec{x} . \lnot \sigma(\vec{x}, \vec{P})$
%If this formula is unsatisfiable, 
%% the candidate solution is in fact a 
%% full solution.
%
If this is the case, then the algorithm terminates.  Otherwise, additional
information is provided to the inductive synthesis phase in the form of a new
counterexample that is added to the {\sc inputs} set and the loop iterates
again.
%% the witness $\vec{x}$ denote an input on which the
%% candidate solution fails to meet the specification.  Thus, $\vec{x}$ is
%% added to the {\sc inputs} set, and the loop iterates again.

It is
worth noting that each iteration of the loop adds a new input to the
finite set $\text{\sc inputs}$ that is used for synthesis.  Given that
the full set of inputs $\mathcal{D}$
%\blue{[which by now should have a name, as suggested to be introduced above]} 
is finite, this means that the refinement loop
can only iterate a finite number of times.

In Sec.~\ref{sec:CEGIS-abstraction-refinement} and 
Sec.~\ref{sec:CEGIS-precision-incrementation}, we explain how we 
specialise the general CEGIS approach in order to perform solution generalisation.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Our synthesis problem}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
The synthesis problem we are trying to solve is the following:
find a digital controller $K$ (see Eq.~\ref{eq:controlaction})
that makes the closed-loop system safe for 
initial state $x_0$, reference signal $r_k$ and input $u_k$
as defined in Sec.~\ref{ssec:safety}. Note that, 
as mentioned at the beginning of the section, 
we consider the reference signal to be 0.
Then, when mapping back to the notation used for describing the general architecture 
of the program synthesizer, the controller $K$ denotes $P$ and 
$(x_0, u_k)$ represents $x$. 
We compute the coefficients for $K$ in the domain $\mathbb{R}<I,F>$.

>>>>

%%%-------------------------------
%%\subsection{Digital implementation of feedback controller}
%%\label{ssec:quantizationerror}
%%%-------------------------------
%%
%%\todo{HERE add a figure similar to Fig. \label{fig:observersystem} to explain digital features of controller.} 
%%
%%\todo{[The remainder of the section needs to be updated. ]}
%%The output and measurable of the plant is being sampled by an ADC, and since the full state of the
%%plant cannot be witnessed directly, the controller software includes an observer that 
%%estimates the state of the plant (see appendix for details on how such an observer
%%is constructed). In this work, we are assuming that all the states measurements are available for the state feedback purpose, and the state observer design may be disregarded.  
%%\todo{[Unclear what the remainder is for: ]}
%%We further presume that there are small tolerances in the system 
%%described in the following form
%%%
%%\begin{align*}
%%\mat{A} \in \mathcal{G}_A=\mat{S}\mat{J}\mat{S}^{-1} : \mat{J} \in \mathcal{G}_J
%%\end{align*}
%%\commentib{What is S?}
%%
%%Where $\mathcal{G}_J$ is a collection of eigenvalue ranges that define the frequency response of the system.
%%We presume that the internal model of the plant is an ARMA model, modified by a
%%state transform that has no tolerances. Tolerances in the eigenvectors will manifest themselves
%%as tolerances in the canonical forms outside of the row/column of interests and complicate the 
%%evaluation. Whilst it is possible to address these with this methodology, we will focus on the case
%%above for simplicity.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{General architecture of the program synthesizer}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\label{synthesizer-general}
The input specification provided to the program synthesizer is of the form
$\exists \vec{P} .  \forall \vec{x}.  \sigma(\vec{x}, \vec{P})$ where
$\vec{P}$ ranges over functions, $\vec{x}$ ranges over ground terms and
$\sigma$ is a quantifier-free formula.  We interpret the ground terms over
some finite domain $\mathcal{D}$.

The design of our synthesizer consists of two phases, an inductive
synthesis phase and a validation phase, which interact via a finite
set of test vectors {\sc inputs} that is updated incrementally.  Given
the aforementioned specification $\sigma$, the inductive synthesis
procedure tries to find an existential witness $\vec{P}$ satisfying
the specification $\sigma(\vec{x}, \vec{P})$ for all $\vec{x}$ in {\sc
  inputs} (as opposed to all $\vec{x} \in \mathcal{D}$).
%
If the synthesis phase succeeds in finding a witness~$\vec{P}$, this
witness is a candidate solution to the full synthesis formula.  We
pass this candidate solution to the validation phase, which checks
whether it is a full solution (i.e., $\vec{P}$ satisfies the
specification $\sigma(\vec{x}, \vec{P})$ for all
$\vec{x}\in\mathcal{D}$).  If this is the case, then the algorithm
terminates.  Otherwise, additional information is provided to the
inductive synthesis phase in the form of a new counterexample that is
added to the {\sc inputs} set and the loop iterates again.

It is worth noting that each iteration of the loop adds a new input to
the finite set $\text{\sc inputs}$ that is used for synthesis.  Given
that the full set of inputs $\mathcal{D}$ is finite, this means that
the refinement loop can only iterate a finite number of times.

\subsection{Solution generalisation} 
Given the large soloution space to be searched during the inductive
synthesis phase, we parametrise the target language, which induces a
lattice of progressively more expressive languages.  This enables us 
to start by attempting to synthesise a program at the lowest point on this lattice
and increase the parameters of the language until we reach a point at
which the synthesis succeeds. Whenever we manage to synthesise
a candidate solution at some lower point on the lattice, 
we apply generalisation to obtain a full solution.

As well as giving us an automatic search
procedure, this parametrisation greatly increases the efficiency of
our system since languages low down the lattice are very easy to
decide safety for (i.e. the validation oracle finds an answer
fast). If a program can be synthesised in a low-complexity language,
the whole procedure finishes much faster than if synthesis had been
attempted in a high-complexity language.

We investigate two different approaches of integrating solution generalisation 
with CEGIS:
\begin{itemize}
\item[1.] The first approach uses abstract acceleration to compute the set of reachable states, which enables it to perform abstraction refinement inside the CEGIS loop (see Sec~\ref{sec:CEGIS-abstraction-refinement}). 
\item[2.] The second approach relies on computing a 
completeness threshold $k$ for each loop such that it is sufficient to 
unroll the corresponding loop $k$ times. For this approach the 
generalisation step is given as an incremental
increase of precision and sampling rate (see Sec.~\ref{sec:CEGIS-precision-incrementation}).
\end{itemize}

% In Sec.~\ref{sec:CEGIS-abstraction-refinement} and 
% Sec.~\ref{sec:CEGIS-precision-incrementation}, we explain how we 
% specialise the general CEGIS approach in order to perform solution generalisation.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Our synthesis problem}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
The synthesis problem we are trying to solve is the following:
find a digital controller $K$ (see Eq.~\ref{eq:controlaction})
that makes the closed-loop system safe for 
initial state $x_0$, reference signal $r_k$ and input $u_k$
as defined in Sec.~\ref{ssec:safety}. Note that, 
we consider the reference signal to be 0 (\todo{refer to wherever this is explained initially}).
Then, when mapping back to the notation used for describing the general architecture 
of the program synthesizer, the controller $K$ denotes $P$ and 
$(x_0, u_k)$ represents $x$. 
We compute the coefficients for $K$ in the domain $\mathbb{R}<I,F>$.
